---
title: "Lab 09 - Grading the professor, Pt. 1"
author: "Hannah Crawley"
date: "3/6/25"
output: github_document
---

## Load Packages and Data

```{r load-packages, message=FALSE}
library(tidyverse) 
library(tidymodels)
library(openintro)
```

```{r}
evals
```


# Part 1: Exploratory Data Analysis

## Exercise 1
> Visualize the distribution of score. Is the distribution skewed? What does that tell you about how students rate courses? Is this what you expected to see? Why, or why not? Include any summary statistics and visualizations you use in your response.

```{r}
ggplot(evals, aes(x = score)) +
  geom_histogram()
```

```{r}
ggplot(evals, aes(x = score)) +
  geom_boxplot()
```

```{r}
summary(evals$score)
```

- Distribution is skewed left meaning that many students are rating courses positively  
- I would expect ratings to be more evenly distributed or perhaps even negative since most students who do reviews are complaining  


```{r exercise1}
# Add your R code here
```

## Exercise 2
> Visualize and describe the relationship between score and the variable bty_avg, a professor’s average beauty rating. See the help page for the function at http://ggplot2.tidyverse.org/reference/index.html.

```{r}
ggplot(evals, aes(x = score, y = bty_avg)) +
  geom_point()
```

- Relationship between beauty and score of professor is slightly positive 

## Exercise 3 
> Replot the scatterplot from Exercise 3, but this time use geom_jitter()? What does “jitter” mean? What was misleading about the initial scatterplot?

```{r exercise3}
ggplot(evals, aes(x = score, y = bty_avg)) +
  geom_jitter()
```
- Jitter adds a small amount of random variation to each point making it easier to see points that potentially overlap  
- this spreads out the points to and makes the distribution look like there is less of a relationship, but still shows skew  
- the first plot is slightly misleading due to the overlap in points 

# Part 2: Linear Regression with a Numerical Predictor 

## Exercise 4 
> Let’s see if the apparent trend in the plot is something more than natural variation. Fit a linear model called m_bty to predict average professor evaluation score by average beauty rating (bty_avg). Based on the regression output, write the linear model.

```{r}
m_bty <- linear_reg() %>%
  set_engine("lm") %>%
  fit(score ~ bty_avg, data = evals)
```

```{r}
summary(m_bty$fit)
```
score(y^) = 3.88 + 0.067(bty_avg)

## Exercise 5 
> Replot your visualization from Exercise 3, and add the regression line to this plot in orange color. Turn off the shading for the uncertainty of the line.

```{r exercise5}
ggplot(evals, aes(x = bty_avg, y = score)) +
  geom_jitter() + 
  geom_smooth(method = "lm", color = "orange", se = FALSE) + 
  labs(title = "Hot or Not?", subtitle = "Teaching Eval v.s. Beauty", x = "Beauty Avg", y = "Eval Score" )
```


## Exercise 6 
> Interpret the slope of the linear model in context of the data.  

Formula: score(y^) = 3.88 + 0.067(bty_avg)

- Slope: for each additional point of beauty average teaching evaluation score is expected to be 0.067 higher on average  


## Exercise 7
>Interpret the intercept of the linear model in context of the data. Comment on whether or not the intercept makes sense in this context. 

- Intercept: For a beauty average of 0 the teaching eval score is expected to be 3.88 -- this intercept does not particularly make since in our context based on the scale presented ... and that the teachers can't be that horrible looking... someone "don't yuck someone else's yum" i guess...

## Exercise 8
> Determine the R^2 of the model and interpret it in context of the data.

```{r}
summary(m_bty$fit)$r.squared

```

- R^2 =  0.03502 -- 3.5% of the variance in the teaching eval score can be explained by avg beauty, not a whole lot 


# Part 3: Linear Regression with a Categorical Predictor 
## Exercise 9
> Fit a new linear model called m_gen to predict average professor evaluation score based on gender of the professor. Based on the regression output, write the linear model and interpret the slope and intercept in context of the data.

```{r}
m_gen <- linear_reg() %>%
  set_engine("lm") %>%
  fit(score ~ gender, data = evals)
```

```{r}
summary(m_gen$fit)
```
> Fit a new linear model called m_gen to predict average professor evaluation score based on gender of the professor. Based on the regression output, write the linear model and interpret the slope and intercept in context of the data.

Equation: score = 4.09 + .14 (gender) where 0 = female and 1 = male   
- intercept = 4.09 -- on average a female professors score is  4.09  
- slope = .14 -- reflects the change in sex from male to female therefore on average the score for male professors is ~ 4.23


## Exercise 10 
> What is the equation of the line corresponding to male professors? What is it for female professors?  

- y(hat) = 4.09 -- female   
- y(hat) = 4.09 + .14 -- male 

## Exercise 11
> Fit a new linear model called m_rank to predict average professor evaluation score based on rank of the professor. Based on the regression output, write the linear model and interpret the slopes and intercept in context of the data.

## Exercise 12
> For Exercise 12, the `relevel()` function can be helpful!Create a new variable called rank_relevel where "tenure track" is the baseline level.

## Exercise 13
> Fit a new linear model called m_rank_relevel to predict average professor evaluation score based on rank_relevel of the professor. This is the new (releveled) variable you created in Exercise 13. Based on the regression output, write the linear model and interpret the slopes and intercept in context of the data. Also determine and interpret the R^2 of the model.

## Exercise 14
> Create another new variable called tenure_eligible that labels "teaching" faculty as "no" and labels "tenure track" and "tenured" faculty as "yes". 

## Exercise 15
> Fit a new linear model called m_tenure_eligible to predict average professor evaluation score based on tenure_eligibleness of the professor. This is the new (regrouped) variable you created in Exercise 15. Based on the regression output, write the linear model and interpret the slopes and intercept in context of the data. Also determine and interpret the R^2 of the model.

