---
title: "Lab 09 - Grading the professor, Pt. 1"
author: "Hannah Crawley"
date: "3/6/25"
output: github_document
---

## Load Packages and Data

```{r load-packages, message=FALSE}
library(tidyverse) 
library(tidymodels)
library(openintro)
```

```{r}
evals
```


# Part 1: Exploratory Data Analysis

## Exercise 1
> Visualize the distribution of score. Is the distribution skewed? What does that tell you about how students rate courses? Is this what you expected to see? Why, or why not? Include any summary statistics and visualizations you use in your response.

```{r}
ggplot(evals, aes(x = score)) +
  geom_histogram()

```

```{r}
ggplot(evals, aes(x = score)) +
  geom_boxplot()
```

```{r}
summary(evals$score)
```

- Distribution is skewed left meaning that many students are rating courses positively  
- I would expect ratings to be more evenly distributed or perhaps even negative since most students who do reviews are complaining  


```{r exercise1}
# Add your R code here
```

## Exercise 2
> Visualize and describe the relationship between score and the variable bty_avg, a professor’s average beauty rating.See the help page for the function at http://ggplot2.tidyverse.org/reference/index.html.

```{r exercise2}
# Add your R code here
```

## Exercise 3 
> Replot the scatterplot from Exercise 3, but this time use geom_jitter()? What does “jitter” mean? What was misleading about the initial scatterplot?

```{r exercise3}
# Add your R code here
```


# Part 2: Linear Regression with a Numerical Predictor 

## Exercise 4 
> Let’s see if the apparent trend in the plot is something more than natural variation. Fit a linear model called m_bty to predict average professor evaluation score by average beauty rating (bty_avg). Based on the regression output, write the linear model.


## Exercise 5 
> Replot your visualization from Exercise 3, and add the regression line to this plot in orange color. Turn off the shading for the uncertainty of the line.

## Exercise 6 
> Interpret the slope of the linear model in context of the data.


## Exercise 7
>Interpret the intercept of the linear model in context of the data. Comment on whether or not the intercept makes sense in this context.

## Exercise 8
> Determine the R^2 of the model and interpret it in context of the data.

# Part 3: Linear Regression with a Categorical Predictor 
## Exercise 9
> Fit a new linear model called m_gen to predict average professor evaluation score based on gender of the professor. Based on the regression output, write the linear model and interpret the slope and intercept in context of the data.

## Exercise 10 
> What is the equation of the line corresponding to male professors? What is it for female professors?


## Exercise 11
> Fit a new linear model called m_rank to predict average professor evaluation score based on rank of the professor. Based on the regression output, write the linear model and interpret the slopes and intercept in context of the data.

## Exercise 12
> For Exercise 12, the `relevel()` function can be helpful!Create a new variable called rank_relevel where "tenure track" is the baseline level.

## Exercise 13
> Fit a new linear model called m_rank_relevel to predict average professor evaluation score based on rank_relevel of the professor. This is the new (releveled) variable you created in Exercise 13. Based on the regression output, write the linear model and interpret the slopes and intercept in context of the data. Also determine and interpret the R^2 of the model.

## Exercise 14
> Create another new variable called tenure_eligible that labels "teaching" faculty as "no" and labels "tenure track" and "tenured" faculty as "yes". 

## Exercise 15
> Fit a new linear model called m_tenure_eligible to predict average professor evaluation score based on tenure_eligibleness of the professor. This is the new (regrouped) variable you created in Exercise 15. Based on the regression output, write the linear model and interpret the slopes and intercept in context of the data. Also determine and interpret the R^2 of the model.

